---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Haley Lam"
date: "03-23-26"
output: github_document
---

### Load packages and data

```{r load}
library(tidyverse)
library(scales)
library(lubridate)
library(zoo)

df <- read_csv("kansas_grouped_rolling_avg.csv")

```

### Exercise 1

Here's some information about the data and the graph:

State: Kansas. 
Timeframe: 2020, July 13 to August 3. 
What's being measured: Average of daily cases per 100k population (i.e., it's adjusted for population size). Two groups: Counties with mask mandates vs. counties without mask mandates.
Source of data: Kansas Department of Health and Environment

### Exercise 2


These counties did not opt out of their mask mandates: Allen, Atchison, Bourbon, Crawford, Dickinson, Douglas, Franklin, Geary, Gove, Harvey, Jewell, Johnson, Mitchell, Montgomery, Morris, Pratt, Reno, Republic, Saline, Scott, Sedgwick, Shawnee, Stanton, and Wyandotte.

"Daily county-level COVID-19 incidence (cases per 100,000 population) was calculated using case and population counts accessed from USAFacts" 
- the data for cases and population came from USAFacts. They provide a link in the footnotes.

### Exercise 3

Load files using webpage URLs
```{r}
cases_raw <- read_csv(
  "https://static.usafacts.org/public/data/covid-19/covid_confirmed_usafacts.csv",
  show_col_types = FALSE
  )
pop_raw <- read_csv(
  "https://static.usafacts.org/public/data/covid-19/covid_county_population_usafacts.csv",
  show_col_types = FALSE
  )

```

```{r}
glimpse(cases_raw)
glimpse(pop_raw)
```

Next step is to figure out how many rows/columns and what they mean, and the issue with wide format data for plotting. Also, to shorten the glimpses.
