---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Haley Lam"
date: "03-23-26"
output: github_document
---

### Load packages and data

```{r load}
library(tidyverse)
library(scales)
library(lubridate)
library(zoo)

df <- read_csv("kansas_grouped_rolling_avg.csv")

```

### Exercise 1

Here's some information about the data and the graph:

State: Kansas. 
Timeframe: 2020, July 13 to August 3. 
What's being measured: Average of daily cases per 100k population (i.e., it's adjusted for population size). Two groups: Counties with mask mandates vs. counties without mask mandates.
Source of data: Kansas Department of Health and Environment

### Exercise 2


These counties did not opt out of their mask mandates: Allen, Atchison, Bourbon, Crawford, Dickinson, Douglas, Franklin, Geary, Gove, Harvey, Jewell, Johnson, Mitchell, Montgomery, Morris, Pratt, Reno, Republic, Saline, Scott, Sedgwick, Shawnee, Stanton, and Wyandotte.

"Daily county-level COVID-19 incidence (cases per 100,000 population) was calculated using case and population counts accessed from USAFacts" 
- the data for cases and population came from USAFacts. They provide a link in the footnotes.

### Exercise 3

Load files using webpage URLs
```{r}
cases_raw <- read_csv(
  "https://static.usafacts.org/public/data/covid-19/covid_confirmed_usafacts.csv",
  show_col_types = FALSE
  )
pop_raw <- read_csv(
  "https://static.usafacts.org/public/data/covid-19/covid_county_population_usafacts.csv",
  show_col_types = FALSE
  )

```

```{r}
glimpse(cases_raw)
glimpse(pop_raw)
```

Wow, this is a giant wall of text. For the "cases_raw" dataset, there are `r nrow(cases_raw)` rows and `r ncol(cases_raw)` columns. Each row is a county and each column represents every single day of the year (starting from 01/22/2020)

For the "pop_raw" dataset, there are `r nrow(pop_raw)` rows and `r ncol(pop_raw)` columns. Each row is a county and the columns have information on the population.

Wide data is a problem for plotting because we cannot make the 'date' column on the x-axis (or any axis). Converting to long format would place all dates in one column.

### Exercise 4

```{r}
cases_clean <- cases_raw %>%
  filter(State == "KS", countyFIPS != 0) %>%
  pivot_longer(
    cols = -c(countyFIPS, `County Name`, State, StateFIPS),  # need to include stateFIPS!
    names_to = "date",
    values_to = "cumulative_cases"
  ) %>%
  select(countyFIPS, `County Name`, date, cumulative_cases) %>%
  mutate(date = lubridate::ymd(date))

pop_clean <- pop_raw %>%
  filter(State == "KS") %>%
  select(countyFIPS, `County Name`, population)
```

### Exercise 5
Here, we restrict to July 5 and August 3.

```{r}
filtered_cases <- cases_clean %>%
  filter(date >= "2020-07-05" & date <= "2020-08-03")
```

### Exercise 6




