---
title: "Lab 07 - Conveying the right message through visualisation"
author: "Haley Lam"
date: "03-23-26"
output: github_document
---

### Load packages and data

```{r load}
library(tidyverse)
library(scales)
library(lubridate)
library(zoo)

df <- read_csv("kansas_grouped_rolling_avg.csv")

```

### Exercise 1

Here's some information about the data and the graph:

State: Kansas. 
Timeframe: 2020, July 13 to August 3. 
What's being measured: Average of daily cases per 100k population (i.e., it's adjusted for population size). Two groups: Counties with mask mandates vs. counties without mask mandates.
Source of data: Kansas Department of Health and Environment

### Exercise 2


These counties did not opt out of their mask mandates: Allen, Atchison, Bourbon, Crawford, Dickinson, Douglas, Franklin, Geary, Gove, Harvey, Jewell, Johnson, Mitchell, Montgomery, Morris, Pratt, Reno, Republic, Saline, Scott, Sedgwick, Shawnee, Stanton, and Wyandotte.

"Daily county-level COVID-19 incidence (cases per 100,000 population) was calculated using case and population counts accessed from USAFacts" 
- the data for cases and population came from USAFacts. They provide a link in the footnotes.

### Exercise 3

Load files using webpage URLs
```{r}
cases_raw <- read_csv(
  "https://static.usafacts.org/public/data/covid-19/covid_confirmed_usafacts.csv",
  show_col_types = FALSE
  )
pop_raw <- read_csv(
  "https://static.usafacts.org/public/data/covid-19/covid_county_population_usafacts.csv",
  show_col_types = FALSE
  )

```

```{r}
glimpse(cases_raw)
glimpse(pop_raw)
```

Wow, this is a giant wall of text. For the "cases_raw" dataset, there are `r nrow(cases_raw)` rows and `r ncol(cases_raw)` columns. Each row is a county and each column represents every single day of the year (starting from 01/22/2020)

For the "pop_raw" dataset, there are `r nrow(pop_raw)` rows and `r ncol(pop_raw)` columns. Each row is a county and the columns have information on the population.

Wide data is a problem for plotting because we cannot make the 'date' column on the x-axis (or any axis). Converting to long format would place all dates in one column.

### Exercise 4

```{r}
cases_clean <- cases_raw %>%
  filter(State == "KS", countyFIPS != 0) %>%
  pivot_longer(
    cols = -c(countyFIPS, `County Name`, State, StateFIPS),  # need to include stateFIPS!
    names_to = "date",
    values_to = "cumulative_cases"
  ) %>%
  select(countyFIPS, `County Name`, date, cumulative_cases) %>%
  mutate(date = lubridate::ymd(date))

pop_clean <- pop_raw %>%
  filter(State == "KS") %>%
  select(countyFIPS, `County Name`, population)
```

### Exercise 5

Here, we restrict to July 5 and August 3.

```{r}
filtered_cases <- cases_clean %>%
  filter(date >= "2020-07-05" & date <= "2020-08-03")
```

```{r}
filtered_cases <- filtered_cases %>%
  group_by(countyFIPS) %>%
  arrange(date) %>%
  mutate(daily_new_cases = cumulative_cases - lag(cumulative_cases)) %>%
  ungroup() %>%
  filter(!is.na(daily_new_cases))
```

### Exercise 6

First, define mandated counties (to later label Mask/No Mask)
```{r}
mandate_counties <- c("Allen", "Atchison", "Bourbon", "Crawford", "Dickinson", 
                      "Douglas", "Franklin", "Geary", "Gove", "Harvey", 
                      "Jewell", "Johnson", "Mitchell", "Montgomery", "Morris", 
                      "Pratt", "Reno", "Republic", "Saline", "Scott", 
                      "Sedgwick", "Shawnee", "Stanton", "Wyandotte")
```

```{r}
filtered_cases <- filtered_cases %>%
  mutate(
    clean_name = str_remove(`County Name`, " County"), 
    mask_mandate = ifelse(clean_name %in% mandate_counties, "Mask", "No Mask")
  )

cases_final <- filtered_cases %>%
  left_join(pop_clean, by = "countyFIPS")
```


### Exercise 7

```{r}
cases_final <- cases_final %>%
  group_by(mask_mandate, date) %>%
  summarize(
    total_new = sum(daily_new_cases, na.rm = TRUE),
    total_pop = sum(population, na.rm = TRUE),
    .groups = "drop"
  )

cases_per100k <- cases_final %>%
  mutate(per_100k = (total_new / total_pop) * 100000)

cases_grouped <- cases_per100k %>%
  group_by(mask_mandate) %>%
  arrange(date) %>%
  mutate(rolling_avg = zoo::rollmean(per_100k, k = 7, fill = NA, align = "right")) %>%
  ungroup() %>%
  filter(date >= "2020-07-12") %>%
  filter(!is.na(rolling_avg)) %>%
  select(date, rolling_avg, mask_mandate)
```

### Exercise 8

This is the misleading graph. The left y-axis labeled "Mask" ranges from 15 to 25. The right y-axis labeled "No Mask" ranges from 4 to 14. This makes the graph uninterpretable, and making the lines visually more alike than different. 
```{r}
misleading_plot <- cases_grouped %>%
  mutate(
    plot_value = case_when(
      mask_mandate == "No Mask" ~ rolling_avg + 11,  # Shift no mask 11 points higher
      TRUE ~ rolling_avg
    )
  ) %>%
  ggplot(aes(x = date, y = plot_value, color = mask_mandate)) +
  geom_line(linewidth = 1.2) + 
  scale_y_continuous(
    name = "Mask Mandate Counties (Cases per 100k)",
    limits = c(15, 25), 
    sec.axis = sec_axis(~ . - 11, name = "No Mask Counties (Cases per 100k)")
  ) +
  labs(title = "Kansas COVID-19 7-Day Rolling Average",
       subtitle = "Misleading Graph!") +
  theme_minimal()

misleading_plot
```

### Exercise 9

Let's do an honest plot.
```{r}
honest_plot <- cases_grouped %>%
  ggplot(aes(x = date, y = rolling_avg, color = mask_mandate)) +
  geom_line(linewidth = 1.2) +
  scale_y_continuous(name = "Daily New Cases per 100k (7-Day Avg)") +
  labs(
    title = "Kansas COVID-19 7-Day Rolling Average",
    subtitle = "Honest Plot With 1 Y-Axis!"
  ) +
  theme_minimal()

honest_plot
```

### Exercise 10
This plot has one y-axis. It does not lead the reader to think that Mask and No Mask counties have similar number of cases (per population). There are differences in baseline rates that were not obvious in the misleading plot.

### Exercise 11
Looking only at the trends, it seems like the No Mask counties don't show an obvious trend, while the Mask counties clearly show decreases in new Covid cases. Since this is in July, the flat line somewhat makes sense. Maybe this is a time when Covid cases started stabalizing? 

### Exercise 12 

The key components to the honest graph. 1) A single y-axis. 2) The axes are scaled approximately to the min and max. 3) Two seaparte lines that are labeled appropriately. 4) The x-axis is interpretable and not crowded.

### Exercise 13-14

One kind of misleading graph would not adjust for the population (i.e., just show raw counts) and does not including rolling avg (i.e., does not take into account the average of the past 7 days)
```{r}
volume_deception_plot <- cases_final %>%
  ggplot(aes(x = date, y = total_new, color = mask_mandate)) +
  geom_line(linewidth = 1.2) +
  labs(
    title = "Kansas COVID-19 Raw Cases",
    y = "Total Daily New Cases (Raw Count)",
    x = "Date",
    color = "Mandate Status"
  ) +
  theme_minimal()

volume_deception_plot
```

As you can see, this plot now makes it look like masks or no masks don't matter too much. It even shows the mask mandated counties have hgiher variability.

